---
title: "Project"
author: "Felix and Renee"
date: "2022-11-20"
output: html_document # need to make this pdf/doc
editor_options: 
  chunk_output_type: console
params:
    printcode: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = params$printcode)
library(psych)
library(RSQLite)
library(knitr)
library(dplyr)
library(lubridate)
library(ggplot2)
```

```{r loadData}
# Load Stations
stations<-read.table('data/ljubljana_station_data_static.csv',sep=',',header=T)
# Load Journeys
journeyConn <- dbConnect(RSQLite::SQLite(), "data/20221120_journey.db")
journeys<-dbGetQuery(journeyConn, "SELECT * FROM Journeys")
```
We loaded data for journeys and stations. We got `r nrow(journeys)` journeys and `r nrow(stations)` stations.
```{r add DayOfTheWeek for everyJourney}
journeys$timestampStartDayOfWeek<-wday(as.Date(journeys$timestampStart), label=TRUE)
journeys$timestampEndDayOfWeek<-wday(as.Date(journeys$timestampEnd), label=TRUE)
```

We added the Days of the Week for each Journey.`r nrow(journeys[journeys$timestampStartDayOfWeek != journeys$timestampEndDayOfWeek,])` Journeys span two days. Thats `r (nrow(journeys[journeys$timestampStartDayOfWeek != journeys$timestampEndDayOfWeek,])/nrow(journeys))*100`%. For further analysis we will therefore only use the starting day of a Journey, since the effect will be minimal.

## 1. What are the most popular stations in Ljubljana?
### all stations

```{r group Journeys by stationStart add Names from stations and sort}
journeys_grouped_stationStart <-
  journeys %>% group_by(stationStart) %>%
  summarise(total_count = n(), .groups = 'rowwise') %>%
  as.data.frame()

journeys_grouped_stationStart <-
  merge(journeys_grouped_stationStart,
        stations[, c('Number', 'Name')],
        by.x = 'stationStart',
        by.y = 'Number')

# Group by totalCount decreasing
journeys_grouped_stationStart <-
  journeys_grouped_stationStart[order(journeys_grouped_stationStart$total_count, decreasing = TRUE), ]
```

```{r plot all start Stations, fig.height = 10}
ggplot(journeys_grouped_stationStart,
       aes(x = reorder(Name,-total_count), y = total_count)) +
  geom_bar(stat = "identity") +
  coord_flip()
```

### a. For starting at ending journeys
### Start stations
```{r group Journeys by stationStart add Names from stations and sort top}
journeys_grouped_stationStart <-
  journeys %>% group_by(stationStart) %>%
  summarise(total_count = n(), .groups = 'rowwise') %>%
  as.data.frame()

journeys_grouped_stationStart <-
  merge(journeys_grouped_stationStart,
        stations[, c('Number', 'Name')],
        by.x = 'stationStart',
        by.y = 'Number')

# Group by totalCount decreasing
journeys_grouped_stationStart <-
  journeys_grouped_stationStart[order(journeys_grouped_stationStart$total_count, decreasing = TRUE), ]
journeys_grouped_stationStart_top10 <-
  journeys_grouped_stationStart[1:10, ]
```


```{r plot top start Stations}
ggplot(journeys_grouped_stationStart_top10,
       aes(x = reorder(Name,-total_count), y = total_count)) +
  geom_bar(stat = "identity") +
  coord_flip()
```

### End stations

```{r group Journeys by stationEnd add Names from stations and sort}
journeys_grouped_stationEnd <- journeys %>% group_by(stationEnd) %>%
  summarise(total_count = n(), .groups = 'rowwise') %>%
  as.data.frame()

journeys_grouped_stationEnd <-
  merge(journeys_grouped_stationEnd,
        stations[, c('Number', 'Name')],
        by.x = 'stationEnd',
        by.y = 'Number')

# Group by totalCount decreasing
journeys_grouped_stationEnd <-
  journeys_grouped_stationEnd[order(journeys_grouped_stationEnd$total_count, decreasing = TRUE), ]
journeys_grouped_stationEnd_top10 <-
  journeys_grouped_stationEnd[1:10, ]
```


```{r plot top end Stations}
ggplot(journeys_grouped_stationEnd_top10,
       aes(x = reorder(Name, -total_count), y = total_count)) +
  geom_bar(stat = "identity") +
  coord_flip()
```

### b. Does this change for the days of the week?

```{r group Journeys by Weekdays and station}
journeys_grouped_weekdays_stationStart <-
  journeys %>% group_by(stationStart, timestampStartDayOfWeek) %>%
  summarise(total_count = n(), .groups = 'rowwise') %>%
  as.data.frame()

journeys_grouped_weekdays_stationStart <-
  merge(
    journeys_grouped_weekdays_stationStart,
    stations[, c('Number', 'Name')],
    by.x = 'stationStart',
    by.y = 'Number'
  )

preferDataFrameByWeekday <- function(dayOfWeek) {
  df <-
    journeys_grouped_weekdays_stationStart[journeys_grouped_weekdays_stationStart$timestampStartDayOfWeek ==
                                             dayOfWeek,]
  df <-
    df[order(df$total_count, decreasing = TRUE), ] # Order by TotalCount
  df <- df[1:10, ] # Get Top 10
  return(df)
}

list_dataframes_by_weekday <- list()
list_dataframes_by_weekday <-
  append(list_dataframes_by_weekday, list(preferDataFrameByWeekday('Mon')))
list_dataframes_by_weekday <-
  append(list_dataframes_by_weekday, list(preferDataFrameByWeekday('Tue')))
list_dataframes_by_weekday <-
  append(list_dataframes_by_weekday, list(preferDataFrameByWeekday('Wed')))
list_dataframes_by_weekday <-
  append(list_dataframes_by_weekday, list(preferDataFrameByWeekday('Thu')))
list_dataframes_by_weekday <-
  append(list_dataframes_by_weekday, list(preferDataFrameByWeekday('Fri')))
list_dataframes_by_weekday <-
  append(list_dataframes_by_weekday, list(preferDataFrameByWeekday('Sat')))
list_dataframes_by_weekday <-
  append(list_dataframes_by_weekday, list(preferDataFrameByWeekday('Sun')))
```

```{r plot top stationsByWeekday}
for(df in list_dataframes_by_weekday){
  plot<-ggplot(df,
       aes(x = reorder(Name, -total_count), y = total_count)) +
  geom_bar(stat = "identity") +
  coord_flip()
  print(plot)
}
```
